---
title: "index"
author: "Parker and Kamiryn"
format: html
editor: visual
---

Our data comes from here: <https://datacatalog.worldbank.org/search/dataset/0038480/Education-Statistics>

## Enrollment Based on Income

```{r}
#| echo: false
library (tidyverse)
library(ggplot2)
library(readxl)
library(here)
url <- "https://databank.worldbank.org/data/download/EdStats_CSV.zip"

dir.create("data")

download.file(
  url, 
  destfile = "data/EdStats_CSV.zip"
)


dest_path <- here("Assignment04/data", "EdStats_CSV.zip")

if (!file.exists(dest_path)){
  download.file(url, 
                destfile = dest_path
  )
}


unzip(
  zipfile = here("Assignment04/data", 
                 "EdStats_CSV.zip"), 
  exdir = here("Assignment04/data")
)

ed_data <- read_csv(
  here("Assignment04/data/", 
       "EdStatsData.csv"
       )
)

primary_ed_data_2000_income <- ed_data %>%
  filter(`Indicator Code`== "SE.PRM.TENR.MA")%>%
  select(`Country Name`,`2000`)

  
colnames(primary_ed_data_2000_income)[2] = "enrollment"

colnames(primary_ed_data_2000_income)[1] = "country"

primary_ed_data_2000_income <- primary_ed_data_2000_income %>%
  filter(country == "High income" | country == "Middle income" | country =="Low income")

primary_ed_data_2000_income %>%
  ggplot(mapping = aes(x=reorder(country,enrollment), y=enrollment)) +
  geom_col(color = "white", fill = "orange") +
  theme_dark()+
  scale_y_reverse()+
  labs(title = "Enrollment Rate (%) in Primary School for Men by Country Income Level",
       subtitle = "Year: 2000", caption = "Source: World Bank")+
  xlab("Income Level") +
  ylab("Enrollment %") 



```

This graph shows the relationship of income class of a country on male matriculation to primary school. This is especially important when looking for indicators of low enrollment in areas where direct enrollment rates may not be readily available. While we can not form any causal assumptions from this data alone, we can determine that extremely high rates of primary enrollment for men is correlated with a country being designated as high income whereas lower rates nearing 50% of enrollment is correlated with a low income designation.\

## Enrollment over time

```{r}
#| echo: false
primary_ed_data_2000_female_enrollment <- ed_data %>%
  filter(`Indicator Code`== "UIS.E.1.G2.F") %>%
  filter(`Country Code` == "HPC") %>%
  select(1, 34:44)

library(reshape2)
library(scales)
primary_ed_data_2000_female_enrollment <- stack(primary_ed_data_2000_female_enrollment)

primary_ed_data_2000_female_enrollment %>%
  filter(ind != "Country Name") %>%
  ggplot(mapping = aes(x=ind, y=values))+
  geom_point(size = 2, shape = "square")+
  geom_smooth()+
  labs(title = "Female Enrollment in Grade 2 over time",
       subtitle = "Years:1999-2009", caption = "Source: World Bank")+
  xlab("Year") +
  ylab("Enrollment number")+
  theme_linedraw()
```

This graph shows female enrollment over time in grade 2 (between 1999 and 2009). This is beneficial because not only do we see an upward trend, we can use this information to glean info about the general way that enrollment moves. We often see missing data in education data, especially the more recent the years are, however, this can help us estimate approximations of what current enrollment may look like if we know any past data point. We can do this by either running a regression or really just finding a line of best fit.

## Gross pre-primary v labor force primary

```{r}
#| echo: false
preprimary_vs_lfprimary <- ed_data %>%
  filter(`Indicator Code`== "SE.PRE.ENRR" | `Indicator Code` == "SL.TLF.BASC.ZS") %>%
  filter(`Country Code` == "USA")

preprimary_vs_lfprimary2 <- t(preprimary_vs_lfprimary)
preprimary_vs_lfprimary2 <- na.omit(preprimary_vs_lfprimary2)

for_graph <- preprimary_vs_lfprimary %>%
  select(34:44)
  
for_graph <- t(for_graph)

for_graph <- na.omit(for_graph)

colnames(for_graph)[2] = "primary_lf" 

colnames(for_graph)[1] = "preprimary"

for_graph2 <- data.frame(
  gross_enrollment_ratio = c(59.92078,59.42370,63.22566, 65.00272, 62.32440, 64.89574, 72.06403),
  lf_with_basic_ed = c(46.114, 47.158, 47.617, 47.675, 47.783, 47.898, 47.891))

for_graph2 %>%
  ggplot(mapping = aes(x= gross_enrollment_ratio, y= lf_with_basic_ed))+
  geom_point(size = 2, shape = "triangle")+
  geom_line()+
  geom_smooth()+
  labs(title = "Gross Enrollment in Pre-Primary v Labor Force with Basic Education",
       subtitle = "Years:1999-2005, USA", caption = "Source: World Bank")+
  xlab("Gross Enrollment Ratio in Pre-Primary (%)") +
  ylab("% of Labor Force with Basic Education")+
  theme_linedraw()

```

This graph is Interesting because it shows a lack of relationship. That is, that regardless of the Gross Enrollment Ratio in Pre-Primary throughout time, there seems to be a consistent % of the labor force with a basic education. This is likely a result of cultural norms or other factors influencing labor force basic education numbers rather than pre-primary. This would make this a terrible indicator to use to account for missing data.

## G

```{r}

primary_ed_data_literacy <- ed_data %>%
  filter(`Indicator Code`== "UIS.ILLPOPF.AG25T64")


primary_ed_data_literacy %>%
  group_by(`Country Code`) %>%
  summarise(n= is.na(`Country Code`)) %>%
  ggplot(
    geom_bar(mapping = aes(x=n))
  )



```

This graph, had it worked properly would show in just one case the sheer number of indicators that are missing from the data. Overall, there is tons of missing data that we have to work with in Public policy and as such, we are not able to simply take data easily and run with it. We need to know whether things can be good indicators, especially as it related to income and literacy rates. How can we best attack the issue of illiteracy when we are unable to know even how bad the issue is? Using indicators is a great way to solve for this.
